---
title: "Homework 2"
author: "Qingcheng Wei"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
sdsd
# Instructions

**Exercises:**  1-3 (Pgs. 49-50); 2-4 (Pg. 51); 2,4 (Pg. 54); 1-4 (Pg. 58)

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining. 

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(nycflights13)
```

# Chapter 3 (Pgs. 49-50)

##  Exercise 1
```{r}
filter(flights,arr_delay>=2)
filter(flights,dest=="IAH"|dest=="Hou")
filter(flights,carrier %in% c("UA","AA","DL"))
filter(flights,month %in% c(7:9))
filter(flights,dep_delay==0&arr_delay>120)
filter(flights,dep_delay>=60&dep_delay-arr_delay>30)
filter(flights,hour>0& hour<=6)
```

##  Exercise 2
it specifies the right and left boundry of the data you are planning to choose.For example, 
```{r}
filter(flights,between(month,7,9))
filter(flights,between(hour,0,6))
```

##  Exercise 3
```{r}
filter(flights,is.na(dep_time))
```
There are a total of 8255 flights have a missing dep_time. It also has missing values of dep_delay, arr_time, arr_delay, and air_time,these rows might be the cancelled flights. 
# Chapter 3 (Pg. 51)

##  Exercise 2
```{r}
arrange(flights, desc(dep_delay))
arrange(flights, year,month,day,dep_time)
```
The most delayed flight is 51 in January 9th
The flight left earliest is 1545 in January 1st
##  Exercise 3
```{r}
arrange(flights,desc(distance/air_time))
```
The fastest flight is 1499 in May 25th
##  Exercise 4
```{r}
arrange(flights,distance)
arrange(flights,desc(distance))
```
Flight 51 travelled the longest 
Flight 3833 travelled the shortest except the one which has not taken off.  
# Chapter 3 (Pg. 54)

##  Exercise 2
It would only show the variable in one column in the tibble.

##  Exercise 4
```{r}
select(flights,contains("TIME",ignore.case=FALSE))

```
By deflaut, select ignores the case difference, by including the argument ignore.case=FALSE, it will make the different cases matter for the coding. 


# Chapter 3 (Pg. 58)

##  Exercise 1
```{r}
mutate(flights,dep_time2=(dep_time%/%100)*60+dep_time%%100,
       sched_dep_time2=(sched_dep_time%/%100)*60+sched_dep_time%%100)
```

##  Exercise 2
```{r}
transmute(flights,air_time,dif=arr_time-dep_time)
transmute(flights,air_time,hour_dif=arr_time%/%100-dep_time%/%100,
          minute_dif=arr_time%%100-dep_time%%100,
          dif=60*hour_dif+minute_dif,
          real_dif=dif-air_time)
?flights
```
I expect the difference would be similar to the air-time. However, it is not the right way to compute the difference, since air_time is calculated in double form for the minutes of flights stayed in air, However, the arr_time and dep_time is an integer number with hour and minute in combined, so the real difference of these two numbers have no real meanings. The first step should switch the time in minutes for variable dep_time and arr_time, then compares. 
##  Exercise 3

```{r}
transmute(flights,dep_delay,sched_dep_time1=sched_dep_time%/%100*60+sched_dep_time%%100,
       dep_time1=dep_time%/%100*60+dep_time%%100,
       dif=dep_time1-sched_dep_time1
       )
```
It's the same, the difference between the actual time to depart with the scheduled time to depart has the same vlaue with delayed time

##  Exercise 4
```{r}

a<-arrange(select(flights,dep_delay,flight),desc(dep_delay))
Rank_dep_delay<-mutate(a,rank=min_rank(desc(dep_delay)))
Rank_dep_delay

```

Luckily, we do not have ties. The top twn delayed flights are flgiht 51, 3535, 3695, 177, 3075,2391,2119,2007,2047, and 172.
