---
title: "Homework 5"
author: "Qingcheng Wei"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  1,3 (Pg. 227); 2 (Pg. 232); 1,2 (Pg. 235); 

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
```

# Chapter 12 (Pg. 227)

##  Exercise 1
```{r}
ggplot(gss_cat,aes(rincome))+geom_bar()+theme(axis.text.x = element_text(angle = 45,hjust=1))+scale_x_discrete(drop=FALSE)
```

##  Exercise 3
```{r}
gss_cat%>%count(relig,denom)

ggplot(gss_cat,aes(x=relig))+geom_bar(aes(fill=denom))+theme(axis.text.x = element_text(angle = 90))
```
It seems denom only applies to the Protestants.

# Chapter 12 (Pg. 232)

##  Exercise 2
```{r}
gss_cat
levels(gss_cat$marital) # it depends. From my view, this sorting is natural. 
levels(gss_cat$race)
gss_cat%>%count(race) # it is principled, based on the count of races from low to high.
levels(gss_cat$rincome) # it is principled, from low wage to high.
levels(gss_cat$partyid) # In my view, it is principled, from republican to independent to democrat.
levels(gss_cat$relig) # it is arbitray
levels(gss_cat$denom) # the same, it is arbitray.
```

As far as I could see from the levels, 
# Chapter 12 (Pg. 235)

##  Exercise 1
```{r}
gss_cat%>%count(partyid)
gss_cat%>% mutate(partyid=fct_collapse(partyid,
  other = c("No answer","Don't know","Other party"),
  Rep = c("Strong republican", "Not str republican","Ind,near rep"),
  Ind = c("Independent"),
  Dem = c("Ind,near dem","Not str democrat","Strong democrat")
))%>% count(year,partyid)%>%
  group_by(year)%>%
  mutate(prop=n/sum(n))%>%
  ggplot(aes(year, prop,color=partyid))+ geom_line()+geom_point()
```

##  Exercise 2
```{r}
gss_cat%>%mutate(rincome=fct_lump(fct_collapse(rincome,Unknown=c("Don't know","Refused","Other","Not applicable","No answer")),n=5))%>%count(rincome)%>%mutate(rincome=fct_recode(rincome,"under 10000"="Other"))

```
I collapses all the unknown values into a single category "Unknown". However, income under 10000 has too many intervals, It could lump into a single category, which by default is named "Other" and I changed it later to "under 10000".