---
title: "Lab 7: Programming"
author: "FIRSTNAME LASTNAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=F)
options(scipen=999)
library(tidyverse)
```

# Introduction

In this lab, you will practice writing functions and using loop functions in R. The loop functions are:

- `lapply()`: Loop over a list and evaluate a function on each element

- `sapply()`: Same as `lapply` but try to simplify the result

- `apply()`: Apply a function over the margins of an array

- `tapply()`: Apply a function over subsets of a vector

- `mapply()`: Multivariate version of `lapply`

# Simulation study
Suppose that $X_1,\ldots,X_n$ are independent and identically distributed (iid) binomial random variables such that
\[
  P(X_i=x\mid k,p)
  ={k\choose x}p^x(1-p)^{k-x},\quad x=0,1,\ldots,k
\]
for all $i=1,\ldots,n$. Assume that both $k$ and $p$ are unknown and use the method of moments to obtain point estimators of both parameters. This somewhat unusual application of the binomial model has
been used to estimate crime rates for crimes that are known to have many unreported
occurrences. For such a crime, both the true reporting rate, $p$, and the total number
of occurrences, $k$, are unknown. Equating the first two sample moments to those of the population yields the system
of equations
\[
  \bar X=kp
  \quad\text{and}\quad
  \frac1n\sum_{i=1}^nX_i^2=kp(1-p)+k^2p^2,
\]
where $\bar X$ is the sample mean. Solving for $k$ and $p$ leads to
\[
  \hat k=\frac{\bar X^2}{\bar X-(1/n)\sum_{i=1}^n(X_i-\bar X)^2}
  \quad\text{and}\quad
  \hat p=\frac{\bar X}{\hat k}.
\]
It is difficult to analyze the performance of $\hat k$ and $\hat p$ analytically so you are asked to perform a simulation study using `R`. The idea is to generate random samples and investigate the performance of $\hat k$ and $\hat p$ using random samples.

### Q1
1. Generate a single simple random sample vector `x` of length `n = 50` from the binomial distribution with the parameters `k = 10`, `p = 0.4`. (1 point)
```{r, eval=T}
k = 10
p = 0.4
x = rbinom(50,k,p)
```

### Q2
2. Write a function that takes a sample vector as its input and returns the estimates of `k` and `p` given above. (4 points)
```{r, eval=T}
est_kp = function(x){
  x_bar = mean(x)
  n = length(x)
  k_hat = x_bar^2/(x_bar-(1/n)*sum((x-x_bar)^2))
  p_hat = x_bar/k_hat
  return(c(k_hat,p_hat))
}
```

### Q3
3. Generate `N = 1000` samples of size `n = 50` (as in the first question) and calculate `N = 1000` estimates of $k$ and `N = 1000` estimates of $p$. Please use Loop functions in this part. `rbinom` and `est_kp` functions are used in this part. (4 points)

```{r, eval=F}
E50 = sapply(lapply(rep(50,1000),rbinom,prob=p,size=k),est_kp)
print(E50)
```

### Q4
4. Repeat Question 3 when `n <- 100` and when `n <- 250`. (2 points)
```{r, eval=T}
E100 = sapply(lapply(rep(100,1000),rbinom,prob=p,size=k),est_kp)
E250 = sapply(lapply(rep(250,1000),rbinom,prob=p,size=k),est_kp)
```

### Q5
5. Estimate the bias and the mean squared error (MSE) of $\hat k$ and the bias and the MSE of $\hat p$ for each sample size (`n <- 50`, `n <- 100` and `n <- 250`). 
a. Do the estimators seem to overestimate or underestimate the parameters? (2 points).
```{r, eval=T}
# bias
rowMeans(E50)-c(10,.4)
rowMeans(E100)-c(10,.4)
rowMeans(E250)-c(10,.4)


# mse
rowMeans(E50-c(10,.4)^2)
rowMeans(E100-c(10,.4)^2)
rowMeans(E250-c(10,.4)^2)
```

b. How do the bias and the MSE change when the sample size increases? (1 point)
ANSWER_HERE: it has no obvious change

### Q6
a. Make a single plot using `ggplot2` that contains three box plots of the estimates of the parameter `k` when `n = 50`, `n = 100`, `n = 250` (the first from the left box plot has to describe the estimates when `n = 50`, the second from the left box plot has to describe the estimates when `n = 100` and the third from the left box plot has to describe the estimates `n = 250`). Include the true value of the parameter as a horizontal line (`geom_hline()` and use the argument `color`) and label the plot appropriately. (3 points)
```{r,eval=T}
df_k<-tibble(
  n=factor(rep(c("50","100","250"),each=1000),c("50","100","250")),
  Estimates=c(E50[1,],E100[1,],E250[1,])
  )
ggplot(data = df_k, mapping = aes(x = n, y = Estimates)) +geom_point()+
	geom_hline(yintercept = k, colour = "red")
```

b. $\hat k$ can obtain values that are far away from the true value of the parameter when the sample size is small and the box plots might not be particularly informative in such a situation. Remove the estimates from the plot that are outside of the interval $[0,50]$ so that the box plots are more informative. (3 points)
```{r, eval=F}
#ggplot(
 # data = filter(df_k, CONDITION), 
 # mapping = aes(x = , y = FILL)
#  ) +
#	FUNCTION +
#	geom_hline(yintercept = k, colour = "red")
```
