---
title: "STOR 320 Tutorial on Modeling II"
author: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=F)
options(scipen=999)
library(tidyverse)    #Essential Functions
library(modelr)       #Helpful Functions in Modeling
library(xtable)

DATA=read_csv("AirWaterTemp.csv",col_types=cols()) #River Data
```

# Introduction

Today, we will work with daily water temperature and air temperature data observed for `r length(unique(DATA$L))` rivers in Spain. The goal of this tutorial is to identify the best model for predicting the maximum water temperature given the maximum air temperature. In the preview below, `W` represents the daily maximum water temperature and `A` represents the daily maximum air temperature. The data contains almost a full year of data for each of the `r length(unique(DATA$L))` different rivers.

```{r,echo=F}
head(DATA)

```


# Part 1: Examining the Relationship

## Chunk 1: Overall Relationship

Please modify the code chunk to visualize the relationship between `A` and `W` with scatterplot (set `alpha=0.3`) and smooth line.
```{r}
DATA%>%ggplot(aes(A,W))+geom_point(alpha=0.3)+geom_smooth(method="lm")+theme_minimal()
```


## Chunk 2: Location-Specific Relationship
Do You Think This Relationship is the Same for All Locations? Please write a function to visualize the relationship between `A` and `W` with scatterplot (set `alpha=0.3`) and smooth line for a specified Location `L`.

```{r,echo=F,eval=F}
WAPLOT.func=function(Location){
  
  DATA%>%filter(L==Location)%>%ggplot(aes(A,W))+geom_point(alpha=0.3)+geom_smooth(method="lm")+theme_minimal()
}
```

Test your function to see if it works:
```{r, eval=F}
WAPLOT.func(103)
WAPLOT.func(105)
WAPLOT.func(918)
```


## Chunk 3: Split Data into Train and Test Sets

```{r}
set.seed(216)
TEST.LOCATIONS=sample(x=unique(DATA$L),size=3,replace=F)


TRAIN = anti_join(DATA,tibble(L=TEST.LOCATIONS),by="L")
TEST = semi_join(DATA,tibble(L=TEST.LOCATIONS),by="L")
```

## Chunk 4: Plots of Relationship for Train and Test Data
Please write a function to visualize the relationship between `A` and `W` with scatterplot (set `alpha=0.3`) and smooth line for a specified Dataset.
```{r,echo=F,eval=T}
WAPLOT2.func=function(DATA){
  DATA%>%ggplot(aes(A,W))+geom_point(alpha=0.3)+geom_smooth(method="lm")+theme_minimal()
}
```

Test the function:
```{r, eval=F}
WAPLOT2.func(TRAIN)
WAPLOT2.func(TEST)
```

# Part 2: Linear Regression Model

## Chunk 1: Fitting Linear Model to Train Data
Please complete the code chunk to fit a linear regression model on the training set. (Formula: `W~A`)
```{r,echo=F,eval=F}
linmod=lm(W~A,data=TRAIN)
summary(linmod)
```

## Chunk 2: Getting Predictions from Linear Model
```{r,eval=F}
TRAIN2 = TRAIN %>% add_predictions(linmod,var="linpred")
TEST2 = TEST %>% add_predictions(linmod,var="linpred")
TRAIN2
TEST2
```

## Chunk 3: Getting Residuals from Linear Model
```{r,eval=F}
TRAIN3 = TRAIN2 %>% add_residuals(linmod,var="linres")
TEST3 = TEST2 %>% add_residuals(linmod,var="linres")
```

# Part 3: Polynomial Regression Model

## Chunk 1: Fitting Polynomial Regression Models
```{r,eval=F}
poly2mod=lm(W~A+I(A^2),data=TRAIN)
poly2mod
poly3mod=lm(W~A+I(A^2)+I(A^3),data=TRAIN)
poly4mod=lm(W~A+I(A^2)+I(A^3)+I(A^4),data=TRAIN)

anova(linmod,poly2mod,poly3mod,poly4mod,test="Chisq")
```

## Chunk 2: Getting Predictions from Polynomial Models

Please add the predictions of polynomial regression models to the training and testing sets with column names `poly2pred`, `poly3pred` and `poly4pred`.
```{r,eval=F}
TRAIN4 =TRAIN3 %>% add_predictions(poly2mod,var="poly2pred") %>% add_predictions(poly3mod,var="poly3pred") %>% add_predictions(poly4mod,var="poly4pred")
TRAIN4
  
TEST4 =TEST3 %>% add_predictions(poly2mod,var="poly2pred") %>% add_predictions(poly3mod,var="poly3pred")%>% add_predictions(poly4mod,var="poly4pred")
```

## Chunk 3: Getting Residuals from Polynomial Models
Please add the residuals of polynomial regression models to the training and testing sets with column names `poly2res`, `poly3res` and `poly4res`.
```{r,eval=F}
TRAIN5 =TRAIN4 %>% add_residuals(poly2mod,var="poly2res")%>% add_residuals(poly3mod,var="poly3res")%>% add_residuals(poly4mod,var="poly4res")
  
  
  

TEST5 =TEST4 %>% add_residuals(poly2mod,var="poly2res")%>% add_residuals(poly3mod,var="poly3res")%>% add_residuals(poly4mod,var="poly4res")
  
  
  
  
```


# Part 4: Nonlinear Logistic Model

## Chunk 1: Logistic Model Investigation
$$y=d+\frac{a}{1+e^{(b-cx)}}$$
```{r,echo=F}
set.seed(216)
EXAMPLE=tibble(
  x=rnorm(10000,mean=0,sd=5),
  y=7+12/(1+exp(4-1*x))
)
ggplot(data=EXAMPLE) +
  geom_point(aes(x=x,y=y)) +
  theme_minimal()
```

## Chunk 2: Essential Functions for Nonlinear Logistic Model
Please write a function to calculate mean squared error for the nonlinear logistic model.
```{r,eval=F}
logistic.model=function(COEF,DATA){
  pred=COEF[1]+COEF[2]/(1+exp(COEF[3]-COEF[4]*DATA$A))
}

MSE.logistic=function(COEF,DATA){
  resi=DATA$W-logistic.model(COEF,DATA)
  mse=mean(resi^2,na.rm=T)
  return(mse)
}
```

## Chunk 3: Estimate Parameters for Nonlinear Logistic Model
```{r,eval=F}
logistic.mod=optim(
  par=c(min(TRAIN$W,na.rm=T),
        max(TRAIN$W,na.rm=T)-min(TRAIN$W,na.rm=T),
        median(TRAIN$A,na.rm=T),
        1),           #Smart Starting Values
  fn=MSE.logistic,    #Function to Minimize
  DATA=TRAIN       #Required Argument
)
print(logistic.mod)

#coef1<-tibble(x=runif(2000,0,50),y=runif(2000,0,50),c=runif(2000,0,50),d=runif(2000,0,50))

# x= apply(coef1,1,MSE.logistic,DATA=TRAIN)
 #y=coef1%>%mutate(rank_x=x,rank_y=rank(x))%>%arrange(rank_y)
# y
 

```

## Chunk 4: Obtain Predictions and Residuals
```{r,eval=F}
TRAIN6=TRAIN5 %>% mutate(logpred=logistic.model(logistic.mod$par,TRAIN5),
                         logres=W-logpred)
TEST6=TEST5 %>% mutate(logpred=logistic.model(logistic.mod$par,TEST5),
                         logres=W-logpred)
TRAIN6
```

# Intermission:

The function `save.image()` in R can be used to save all objects in the global environment. This is very helpful when you want work off your results without rerunning all previous R code. The name of the exported information should contain the file extension *.Rdata*. These files can be extremely large depending how much RAM was utilized in your R session.  The function `load()` can be used to import a previous workspace.

For more information on *.Rdata* file types, see https://fileinfo.com/extension/rdata for help.

```{r,eval=F}
save.image("Tutorial.Rdata")
```

```{r, eval=F}
load("PATH_TO_THE_FILE")
```


# Part 5: Evaluation by Visualization

## Chunk 1: Plotting Models
```{r,echo=F,eval=F}
TEST6 %>%
  select(L,A,W,linpred,poly2pred,poly3pred,poly4pred,logpred)%>%
  gather(linpred:logpred,key="Model",value="Pred",factor_key=T) %>%
    ggplot() + 
      geom_point(aes(x=A,y=W),color="gray") + 
      theme_minimal() +
      geom_line(aes(x=A,y=Pred,color=Model),size=2)
```

## Chunk 2: Predicted Versus Actual Max Water Temperature
```{r,echo=F,eval=F}
TEST6 %>%
  select(L,A,W,linpred,poly2pred,poly3pred,poly4pred,logpred)%>%
  gather(linpred:logpred,key="Model",value="Pred",factor_key=T) %>%
    ggplot() + 
    geom_point(aes(x=W,y=Pred,color=Model)) + 
    geom_abline(a=0,b=1,color="black",size=2) +
    xlab("Maximum Water Temperature") +
    ylab("Predicted Water Temperature") +
    theme_minimal()
```

## Chunk 3: Plotting Residuals Versus Time
```{r,echo=F,eval=F}
TEST6 %>%
  select(L,A,W,TIME,linres,poly2res,poly3res,poly4res,logres)%>%
  gather(linres:logres,key="Model",value="Res",factor_key=T) %>%
    ggplot() + 
      geom_line(aes(x=TIME,y=Res),color="lightskyblue2") + 
      geom_hline(yintercept=0,color="black",linetype=2,size=1.5) +
      xlab("Time") +
      ylab("Residual") +
      theme_dark() +
      facet_grid(Model~.)
```

## Chunk 4: Evaluating The Location-Specific Error
```{r,echo=F,eval=F}
TEST6 %>%
  select(L,A,W,TIME,linpred,logpred) %>%
  gather(linpred:logpred,key="Model",value="Pred",factor_key=T) %>%
    ggplot() +
      geom_point(aes(x=A,y=W),alpha=0.2) +
      geom_line(aes(x=A,y=Pred,color=Model),size=2) +
      theme_minimal()+facet_grid(L~.)

TEST6 %>%
  select(L,A,W,TIME,linres,logres) %>%
  gather(linres:logres,key="Model",value="Res",factor_key=T) %>%
  ggplot() +
  geom_point(aes(x=A,y=Res,color=Model)) +
  geom_hline(yintercept=0) +
  theme_minimal()+facet_grid(L~.)

TEST6 %>%
  select(L,A,W,TIME,linres,logres) %>%
  gather(linres:logres,key="Model",value="Res",factor_key=T) %>%
  ggplot() +
  geom_point(aes(x=TIME,y=Res,color=Model)) +
  geom_hline(yintercept=0) +
  theme_minimal()+facet_grid(L~.)
```


# Part 6: Evaluation by Numerical Summary

## Chunk 1: Functions Required For Evaluating Prediction
```{r,eval=F}
bias.func=function(res){
  bias=mean(res,na.rm=T)
  return(bias)
}

mae.func=function(res){
  mae=mean(abs(res),na.rm=T)
  return(mae)
}

rmse.func=function(res){
  mse=mean(res^2,na.rm=T)
  rmse=sqrt(mse)
  return(rmse)
}
```

## Chunk 2: Checking Functions 
Please use `apply` function to calculate Bias, MAE and rMSE for the models.
```{r,eval=F}
ex.res.mat=TEST6 %>% select(linres,poly2res,poly3res,poly4res,logres)
apply(COMPLETE)
apply(COMPLETE)
apply(COMPLETE)
```

## Chunk 3: Get Table Quickly
```{r,echo=F,eval=F}
SUMM1=TEST6 %>%
  select(L,A,W,TIME,linres,poly2res,poly3res,poly4res,logres)%>%
  rename(Linear=linres,`Poly(2)`=poly2res,`Poly(3)`=poly3res,`Poly(4)`=poly4res,Logistic=logres)%>%
  gather(Linear:Logistic,key="Model",value="Residual",factor_key=T)
SUMM2= SUMM1 %>% 
  group_by(Model) %>%
  summarize(MB=bias.func(Residual),
            MAE=mae.func(Residual),
            RMSE=rmse.func(Residual))
print(SUMM2)
```

## Chunk 4: HTML Formatted Table

```{r,echo=F,eval=F}
SUMM3=xtable(SUMM2,digits=4,align=c("l","l","r","r","r"))
print(SUMM3,include.rownames=F,type="html")
```
















