---
title: "FIFA Qingcheng"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(generics)
library(modelr)       #Helpful Functions in Modeling
library(xtable)
```

```{r}
table<-read.csv("FIFA_20.csv")
table$work_rate=factor(table$work_rate,levels=c("Low","Medium","High"))
levels(table$work_rate)
table
```
```{r}

table2<-table%>%select("name","age":"sliding_tackle")
print(table2)
#%>%crossv_kfold(10)
mod<-function(i){
  mod1=lm(composure~.,data=table2)
}
result<-table2%>%mutate(tr.model=map(train,mod))%>%
  mutate(predict=map2(test,tr.model,~argument(.y,newdata=.x)))%>%
  select(predict)%>%unnest()
mse<-function(x,y){
  mse=mean(abs(x-y))
  return(mse)
}
pred<-mse(result$.fitted,result$composure)

```

```{r}
low_wage=quantile(table$wage_eur,0.3333)
middle_wage=quantile(table$wage_eur,0.6666)

wage=rep(NA,length(table$wage_eur))
for (i in seq_along(table$wage_eur)){
  if (table$wage_eur[i]<=low_wage){wage[i]="Low Wage"}else if 
 (low_wage<table$wage_eur[i]&table$wage_eur[i]<=middle_wage){wage[i]="Medium Wage"} else
    {wage[i]="High Wage"}
}
table2<-table%>%mutate(wage_eur=wage)
table2$work_rate<-factor(table2$work_rate,levels=c("Low","Medium","High"),ordered=TRUE)
table2$wage_eur<-factor(table2$wage_eur,levels=c("Low Wage","Medium Wage","High Wage"))
```

Q1: What is the distribtion of the shooting scores among these players across different work rates? And how it will change across different wage levels?

```{r}
chart1<-table2%>%
  ggplot(aes(reorder(work_rate,shooting,median),shooting))+geom_boxplot()+facet_wrap(~wage_eur)+xlab("Work Rate")
chart1

```

Q1: What is the distribtion of the defending scores among these players across different work rates. And how it will change across different wage levels.
```{r}
chart2<-table2%>%ggplot(aes(reorder(work_rate,defending,median),defending))+geom_boxplot()+facet_wrap(~wage_eur)+xlab("Work Rate")
chart2
```
Follow-up: How is the income distributes among players with low work rate and high work rate. Which type of players(determined by the shooting and defending score) have higher preferences(determined by the average wages) in this distance ran. 

```{r}
#range(table2$shooting)
#range(table2$defending)
#table%>%ggplot(aes(shooting))+geom_histogram()
ces<-function(x){
  y = rep(NA, length(x))
  for (i in seq_along(x)){
    if (x[i]>10&x[i]<30){
      y[i]="10-30"
    }else if(x[i]>30&x[i]<50){
      y[i]="30-50"
    }else if(x[i]>50&x[i]<60){
      y[i]="50-60"
    }else if(x[i]>60&x[i]<70){
      y[i]="60-70"
  }else if(x[i]>70&x[i]<94){
      y[i]="70-93"
  }} 
  return(y)
}
plot1<-table%>%filter(work_rate=="High",!is.na(shooting)&!is.na(defending))%>%mutate(shooting=ces(shooting),defending=ces(defending))%>%group_by(shooting,defending)%>%mutate(meanIncome=mean(wage_eur,na.rm=T))%>%ggplot(aes(shooting,defending))+geom_tile(aes(fill=meanIncome))

plot1
plot2<-table%>%filter(work_rate=="Low")%>%mutate(shooting=ces(shooting),defending=ces(defending))%>%group_by(shooting,defending)%>%mutate(meanIncome=mean(wage_eur,na.rm=T))%>%ggplot(aes(shooting,defending))+geom_tile(aes(fill=meanIncome))

plot2

#ggplot(table,aes(defending))+geom_histogram(aes(fill=work_rate))
#ggplot(table,aes(shooting))+geom_histogram(aes(fill=work_rate))
```
Work rate typically represents the distance travelled by a player in a game. 
Some sports reporters claim work rate is one of the most important determinators for a good soccer player. That is not simply true, by the first round of investigation, we could see the players with lower work rate have higher defending scores, which mean that their role is mainly for defending and high work rate players are typically the attackers with higher shooting scores. This means that players who are majorly defenders travel less distance whereas attackers and shooters need to constantly move so that their work rate is much higher. Also, the median and overall shooting and defending score is increasing with higher wage levels. That means wage could ideally reflect the ability or value of a player in a game no matter it is a defender or attacker.  


So in order to justify this guess and see which type of players, defender or shooter, has more values by comparing their wages in separate graph for high and low work rate.I divided shooting and defending scores into intervals in which each of them shall have similar number of players. As a result, We could see the values for high work rate players are almost twice higher than the low work rate players, which means attackers or shooters brings higher values to a team or they have higher market force. Moreover, in high work-rate players, when defending score interval keeps constant, the average wages increase as shooting score increases. But when we look at the shooting interval 70-93, the average wages does not increase as defending score increases. For instance, even 50-60 players have higher average wages than 70-93's.

By looking at the second graph, low work-rate players have much less than high work-rate players. The missing blocks represent no players fall into these criteria. In this graph, conversely, we could see that holding the defending score in 70-93 interval constant, the average wages do not show a increasing pattern as shooting score increases, which players'wages with shooting score 50-60 is even higher than the 60-70 players'wages. 

Overall, the first graph shows that players with high work-rate is typically a shooter or attacker whose shooting score is determinant of their wages regardless their defending levels. Conversely, for players who have low work-rate are typically defenders whose defending score is most important factor. 
```{r}
#Com=function(variable1){
t1=tibble()
for (i in 1:3){
a<-table%>%filter(work_rate==levels(work_rate)[i])
b<-lm(wage_eur~shooting,data=na.omit(a))
t1=rbind(t1,tidy(b))
}


t2=tibble()
for (i in 1:3){
a<-table%>%filter(work_rate==levels(work_rate)[i])
b<-lm(wage_eur~defending,data=a)
table<-table%>%filter(work_rate==levels(work_rate)[i])%>%add_predictions(b,var="pred3")
t2=rbind(t2,tidy(b))
}

```



```{r}
#pred<-lm(wage_eur~shooting+work_rate,data=table)
#table<-table%>%add_predictions(pred,var="pred2")
#table%>%ggplot()+geom_point(aes(x=shooting,y=wage_eur,color=work_rate))+geom_line(aes(x=shooting,y=pred2,color=work_rate))

table2<-table%>%select("age":"sliding_tackle")%>%crossv_kfold(10)
mod1<-lm(composure~)



