---
title: "Lab 10: Categorical Variables"
author: "FIRSTNAME LASTNAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=F)
options(scipen=999)
library(tidyverse)
library(modelr)
library(broom)
```

# Introduction

In this lab, you will build predictive models for board game ratings. The dataset below was scraped from [boardgamegeek.com](www.boardgamegeek.com) and contains information on the top 4,999 board games. We start by creating the following 8 new variables:

- *duration=2018-year+1*
- *vote.per.year=num_votes/duration*
- *own.per.year=owned/duration*
- *player.range=max_players-min_players*
- *log_vote=log(num_votes+1)*
- *log_own=log(owned+1)*
- *diff_rating=avg_rating-geek_rating*

The table below gives a preview of the current data.

```{r}
bgg<-read.csv("bgg.csv")
bgg2=bgg[,c(4:13,15:20)]

bgg3=bgg2 %>%
  mutate(duration=2018-year+1,
         vote.per.year=num_votes/duration,
         own.per.year=owned/duration,
         player.range=max_players-min_players,
         time.range=max_time-min_time,
         log_vote=log(num_votes+1),
         log_own=log(owned+1),
         diff_rating=avg_rating-geek_rating)
head(bgg3)

```


# Extended Board Game Analysis

### Q1

My favorite mechanics in bard games are "Co-operative Play", "Tile Placement", "Worker Placement", and "Card Drafting". I want you to create the three following binary variables:

- *coop = 1 if Co-operative Play is a mechanic in the game and 0 otherwise*
- *tile = 1 if Tile Placement is a mechanic in the game and 0 otherwise*
- *worker = 1 if Worker Placement is a mechanic in the game and 0 otherwise*
- *draft = 1 if Card Drafting is a mechanic in the game and 0 otherwise*

Notice how we can use the `str_detect()` function to return TRUE if a pattern exists in a string, and then, we use `as.numeric()` to convert TRUE to a 1. We get 0 whenever the pattern is not detected in the string. This function can be vectorized like I do below in the creation of coop. Repeat this for the other variables that you are asked to create. Put this all in a new object named `bgg4`.

Look at random observations in the data to make sure everything worked.

```{r,eval=T}
x=c("Hello","Little","Buddy")
as.numeric(str_detect(x,pattern="Buddy"))
as.numeric(str_detect(x,pattern="Friend"))

bgg4=bgg3 %>%
  mutate(coop=as.numeric(str_detect(mechanic,pattern="Co-operative Play")),
         tile=as.numeric(str_detect(mechanic,pattern="Tile Placement")),
         worker=as.numeric(str_detect(mechanic,pattern="Worker Placement")),
         draft=as.numeric(str_detect(mechanic,pattern="Card Drafting")))
```

### Q2

We want to explore the relationship with these newly created categorical (binary) variables and the geek rating. Build a linear regression model using the 4 variables created above to predict geek rating. Save the model as an object called `mod1` and print out the model using the `tidy()` function. 

```{r,eval=T}
mod1=lm(geek_rating~coop+tile+worker+draft,data=bgg4)
tidy(mod1)
```

Question: In a complete sentence, interpret the intercept of the model. Reference the estimate of the intercept and explain what it represents in way a common person would understand.

the estimation of geek_rating, if a game is not co_operative play, contains tile placement,worker placement and card drafting. 



Question: In a complete sentence, explain the estimated coefficient for the Worker Placement variable. Reference the estimated value and explain in a way that a common person would understand.

The coefficent is how much could the geek_rating increases if a games has the following staffs. 


### Q3

Use the `data_grid()` function to create an object called `GRID` that contains all combinations of four categorical (binary) variables. Then, use the predict function to get the predicting geek ratings for each combination of these four variables. 

In the end, use the `head()` function to print out the grid of fitted values along with their confidence intervals. 

```{r,eval=T}
GRID=bgg4 %>%
        data_grid(
          coop=coop,
          tile=tile,
          worker=worker,
          draft=draft
        )

GRID2=cbind(GRID,predict(mod1,newdata=GRID,interval="confidence"))

head(GRID2)

```



### Q4

From the last model, it seems that Worker Placement and Card Drafting games typically having higher geek ratings than games without these mechanics. Create a model called `mod2` that only has these two variables in the model along with the interaction term between these two variables. Use the `tidy()` function to display the model.

```{r,eval=T}
mod2=lm(geek_rating~worker+draft+worker*draft,data=bgg4)
tidy(mod2)
```




### Q5

Use the `data_grid()` function as in Q3 for this model named `mod2`, and then after obtaining the fitted values and confidence intervals, plot the confidence intervals as seen in lecture. Place `worker` on the x-axis and use `draft` to modify the color. Find a way to change the values of "0" and "1" to "No" and "Yes". I recommend using the `ifelse()` function.

```{r,eval=T}

GRID3=bgg4 %>%
        data_grid(
          worker=unique(worker),
          draft=unique(draft)
        )

GRID4=cbind(GRID3,predict(mod2,newdata=GRID3,interval="confidence"))%>%mutate(draft=ifelse(draft==0,"No","Yes"))

ggplot(GRID4) +
  geom_pointrange(aes(x=worker,y=fit,ymin=lwr,ymax=upr,color=draft),
                  position=position_dodge(width=0.5),size=1.5) +
  theme_minimal()+theme(text=element_text(size=20))

```

Question: In complete sentences, why is the confidence interval for board games where there is no worker placement and no card drafting so small that only a dot appears in the picture?

because it will match with the interception point, coefficent will not take place in this question. 




### Q6

Below I create a dataset called `final.bgg` that removes variables that were used to create other variables and removes some other categorical variables.

```{r,eval=T}
final.bgg=bgg4 %>%
  select(-max_players,-max_time,-year,-avg_rating,-num_votes,-owned,-category,-mechanic,-designer,-names)

head(final.bgg)
```

Create a new dataset called `final.bgg2` where you create a new variable called `Favorite` that equals 1 if the the board game has at least one of my four favorite mechanics, and then remove the four variables we created named `coop`, `tile`, `worker`, and `draft`. Then use the `str()` function to show `final.bgg2`

```{r,eval=T}
final.bgg2=final.bgg %>%
              mutate(Favorite=ifelse((coop+tile+worker+draft)>=1,1,0)) %>%
              select(-coop,-tile,-worker,-draft)
str(final.bgg2)
Favorite=1

mean(final.bgg2$Favorite>=1)
```

Question: In a complete sentence, what percent of games in `final.bgg` have at least one my four favorite mechanics. Use inline R code to insert your answer directly into your sentence. You can calculate this directly from the new variable named `Favorite`. Use the round function to round your percentage to 2 decimal places.
'r sum(final.bgg2$Favorite)/4999'


### Q7

Build 3 different logistic regression models called "Model_1", "Model_2" and "Model 3" to classify a game as a favorite of Dr. Mario. Each of the three models should have five different variables in them. You can pick from any of the variables in `final.bgg2`. You can have variables that are in multiple models, but none of the three models should have the same exact 5 variables. 

Then, I want you to build a table that shows the name of the models and the proportion of time that the model accurately classified a board game as one of my favorites. You can calculate this measure of accuracy in the original dataset `final.bgg2`. There is no need here to split the data into training and testing datasets. Then, print out this entire table.

```{r,eval=T}
Model_1=glm(Favorite~min_players+avg_time+min_time+geek_rating+age,data=final.bgg2,family="binomial")
Model_2=glm(Favorite~min_players+avg_time+min_time+geek_rating+weight,data=final.bgg2,family="binomial")
Model_3=glm(Favorite~min_players+avg_time+min_time+geek_rating+duration,data=final.bgg2,family="binomial")


final.bgg3=final.bgg2 %>%
              mutate(p1=predict(Model_1,newdata=final.bgg2),
                     p2=predict(Model_2,newdata=final.bgg2),
                     p3=predict(Model_3,newdata=final.bgg2),
                     S1=ifelse(p1<0.5,0,1),
                     S2=ifelse(p2<0.5,0,1),
                     S3=ifelse(p3<0.5,0,1))

Model_1.accuracy=mean(final.bgg3$Favorite==final.bgg3$S1)
Model_2.accuracy=mean(final.bgg3$Favorite==final.bgg3$S2)
Model_3.accuracy=mean(final.bgg3$Favorite==final.bgg3$S3)

tibble(Model=c("Model_1","Model_2","Model_3"),Accuracy=c(Model_1.accuracy,Model_2.accuracy,Model_3.accuracy))

              
```

Question: In complete sentences, describe your best model and how accurately it classified board games as a favorite of mine. Talk about what variables are in your best model. Talk about how good the best model was relative to the other 3 models.
The best one is Model2. They are basically 


